Prospects of detecting gamma-ray emission from galaxy clusters: cosmic rays and dark matter annihilations

### 2011-04-05 paper.tex
new fig

Even though the bulk of substructures have been erased in the central
regions of DM halos, a cluster in projection has a significant
enhancement due to substructures at a radius of just a few percent of
$\rvir$. In Fig.~\ref{fig:radial_emis} we show the luminosity
dominated radius for substructures and smooth density profiles, both
for three different mass scales. It clearly shows how that the main
contribution from the substructures is coming from outer parts of the
DM halos, while the main contribution from the smooth density profiles
peaks around the scale radius. SHOW WHERE THE FLUX PEEKS WITH EQUATION
SEC DERIV
\begin{figure}
  \includegraphics[width=0.99\columnwidth]{figures/emissiv.sub.eps}
  \caption{\it Luminosity dominated radius for different mass
    scales. The solid lines show the emissivity from a cuspy
    $\beta=1.0$ NFW density profile for three different mass scales;
    an $\mvir=10^{15}\,\msun$ galaxy cluster (red), an
    $\mvir=10^{12}\,\msun$ galaxy (orange), and an
    $\mvir=10^{8}\,\msun$ dwarf galaxy (green). The dashed lines show
    the emissivity from substructures for the same three mass
    scales. All emissivities have been normalized with the luminosity
    within $\rvir$ from a cuspy NFW profile. We have assumed the
    standard value for the limiting substructure mass of
    $M_\rmn{lim}=10^{-6}\,\msun$.}
  \label{fig:radial_emis}
\end{figure}

@@

The normalization constant for the dust $N_\dust(\lx)$ is derived in the
following four steps: (1) We use a cluster with the the average mass
$\mvir=4.0\times10^{13}\msun$ and estimate the X-ray luminosity from
the X-ray to mass scaling relations derived for the Representative
XMM-Newton Cluster Structure Survey (REXCESS) in
\cite{2009A&A...498..361P}:
\begin{equation}
\lx=1.18\times 10^{44}\,\rmn{erg\,s}^{-1}
\left(\frac{M_{200}}{10^{14}\msun}\right)^{1.81}\,,
\end{equation}
where we have used that $M_\rmn{200} \approx 1.58\,M_\rmn{500}$ for a
small cluster \footnote{We derive the factor 1.58 using
  $M_\rmn{200}/M_\rmn{500}$ for the clusters in the HIFLUGCS catalogue
  with a mass similar to $\mvir=4.0\times10^{13}\msun$ that we use to
  derive the luminosity from dust.}. (2) From the IR to X-ray scaling
relation in Eq.~(\ref{eq:SD-X}) we can now estimate calculate the
luminosity from the dust for a $\mvir=4.0\times10^{13}\msun$ cluster;
$L_\dust=4.7\times10^{44}\,\rmn{erg\,s}^{-1}$. (3) We generalize
Eq.~(\ref{eq:L_star}) to an arbitrary cluster mass using
Eq.~(\ref{eq:SD-X}) together with the cluster X-ray luminosities in
the extended HIFLUGCS catalogue. (4) The normalization constant for
the dust can now be determined from from the above steps together with
Eq.~(\ref{eq:E_SD}) where we integrate over the cluster volume and
spectral distribution of the dust. It is given by
\begin{equation}
 N_\dust(\lx) = 1.4\times10^{-9}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,.
\label{eq:N_stars}
\end{equation}
Note that since the luminosities include the contribution from
galaxies, we have integrated the spatial distribution including
galaxies to derive the normalization. Once the normalization is fixed,
the SD energy densities are derived from the spatial distribution
where galaxies have been excluded.

@@

\smallskip
We wish to thank...

>>>>

In Fig.~\ref{fig:radial_emis} we show the radius with the maximum
luminosity for substructures and smooth density profiles, both for
three different mass scales. We find that the smooth luminosity per
logaritmic radial bin peaks at $x \dd^2 L/\dd x^2+\dd L/\dd x == 0
\rightarrow r \propto \rvir/3c$. Hence the main contribition to the
luminosity for large mass scales, which are usually assosiated with a
low concentrated DM halos, is focused to the regime around 10\% of
$\rvir$.  It clearly shows how that the main contribution from the
substructures is coming from outer parts of the DM halos, while the
main contribution from the smooth density profiles peaks around the
scale radius. Not that, even though the bulk of substructures have
been erased in the central regions of DM halos, a cluster in
projection has a significant enhancement due to substructures at a
radius of just a few percent of $\rvir$.
\begin{figure}
  \includegraphics[width=0.99\columnwidth]{figures/emissiv.sub.eps}
  \caption{\it Luminosity dominated radius for different mass
    scales. The solid lines show the emissivity from a cuspy
    $\beta=1.0$ NFW density profile for three different mass scales;
    an $\mvir=10^{15}\,\msun$ galaxy cluster (red), an
    $\mvir=10^{12}\,\msun$ galaxy (orange), and an
    $\mvir=10^{8}\,\msun$ dwarf galaxy (green). The dashed lines show
    the emissivity from substructures for the same three mass
    scales. All emissivities have been normalized with the luminosity
    within $\rvir$ from a cuspy NFW profile. We have assumed the
    standard value for the limiting substructure mass of
    $M_\rmn{lim}=10^{-6}\,\msun$.}
  \label{fig:radial_emis}
\end{figure}

@@

The normalization constant for the dust $N_\dust(\lx)$ is derived in the
following four steps: (1) We use a cluster with the the average mass
$\mvir=4.0\times10^{13}\msun$ and estimate the X-ray luminosity from
the X-ray to mass scaling relations derived for the Representative
XMM-Newton Cluster Structure Survey (REXCESS) in
\cite{2009A&A...498..361P}:
\begin{equation}
\lx=1.18\times 10^{44}\,\rmn{erg\,s}^{-1}
\left(\frac{M_{200}}{10^{14}\msun}\right)^{1.81}\,,
\end{equation}
where we have used that $M_\rmn{200} \approx 1.58\,M_\rmn{500}$ for a
small cluster \footnote{We derive the factor 1.58 using
  $M_\rmn{200}/M_\rmn{500}$ for the clusters in the HIFLUGCS catalogue
  with a mass similar to $\mvir=4.0\times10^{13}\msun$ that we use to
  derive the luminosity from dust.}. (2) From the IR to X-ray scaling
relation in Eq.~(\ref{eq:SD-X}) we can now estimate calculate the
luminosity from the dust for a $\mvir=4.0\times10^{13}\msun$ cluster;
$L_\dust=4.7\times10^{44}\,\rmn{erg\,s}^{-1}$. (3) We generalize
Eq.~(\ref{eq:L_star}) to an arbitrary cluster mass using
Eq.~(\ref{eq:SD-X}) together with the cluster X-ray luminosities in
the extended HIFLUGCS catalogue. (4) The normalization constant for
the dust can now be determined from from the above steps together with
Eq.~(\ref{eq:E_SD}) where we integrate over the cluster volume and
spectral distribution of the dust. It is given by
\begin{equation}
 N_\dust(\lx) = 1.4\times10^{-9}\,
\left(\frac{\lx}{10^{44}\rmn{erg\,s}^{-1}}\right)^{0.53}\,.
\label{eq:N_dust}
\end{equation}
Note that since the luminosities include the contribution from
galaxies, we have integrated the spatial distribution including
galaxies to derive the normalization. Once the normalization is fixed,
the SD energy densities are derived from the spatial distribution
where galaxies have been excluded.

@@

\smallskip
Anders Pinzke acknowledges NSF grant AST 0908480 for support.
